---
title: "CAUV Projections, 2020 and beyond"
author: Robert Dinterman
date: "2020-06-25"
output:
  revealjs::revealjs_presentation:
    theme: night
    highlight: pygments
    center: false
    css: "4-recommendations.css"
    self_contained: true
    fig_width: 7
    fig_height: 5
    reveal_options:
      slideNumber: true
      previewLinks: true
  beamer_presentation: 
  powerpoint_presentation: default
---

```{r, include=FALSE}
library(knitr)
opts_knit$set(root.dir = normalizePath("../"))
read_chunk("../3-proj/3-projection-tidy.R")
```

```{r global_options, include=FALSE}
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, out.width = "100%", fig.pos = "H")
```

```{r start, include=FALSE}
```

```{r expected, include=FALSE}
```

```{r low, include=FALSE}
```

```{r high, include=FALSE}
```

```{r projections, include=FALSE}
```

# Key Findings

## Highlights

- Current projection for average CAUV value in 2020 is `r dollar(ohio_soils_all$expected[ohio_soils_all$year == 2020], accuracy = 5)` which would represent a change of `r percent((ohio_soils_all$expected[ohio_soils_all$year == 2020] - ohio_soils_all$expected[ohio_soils_all$year == 2019]) / ohio_soils_all$expected[ohio_soils_all$year == 2019], accuracy = 1)` from the previous tax year. This is the first year without the "phase-in".
    - From 2017-2020 a `r percent((ohio_soils_all$expected[ohio_soils_all$year == 2020] - ohio_soils_all$expected[ohio_soils_all$year == 2017]) / ohio_soils_all$expected[ohio_soils_all$year == 2017], accuracy = 1)` decrease in average CAUV values for counties receiving updates to their property taxes.
- The projection for the average CAUV value in 2021 is `r dollar(ohio_soils_all$expected[ohio_soils_all$year == 2021], accuracy = 5)`. The average CAUV value projections could end up as low as `r dollar(ohio_soils_all$low[ohio_soils_all$year == 2021], accuracy = 5)` or as high as `r dollar(ohio_soils_all$high[ohio_soils_all$year == 2021], accuracy = 5)`.
    - From 2018-2021 a `r percent((ohio_soils_all$expected[ohio_soils_all$year == 2021] - ohio_soils_all$expected[ohio_soils_all$year == 2018]) / ohio_soils_all$expected[ohio_soils_all$year == 2018], accuracy = 1)` change from the previous triennial update.
- The very early projection for the average CAUV value in 2022 is `r dollar(ohio_soils_all$expected[ohio_soils_all$year == 2022], accuracy = 5)`. The average CAUV value projections could end up as low as `r dollar(ohio_soils_all$low[ohio_soils_all$year == 2022], accuracy = 5)` or as high as `r dollar(ohio_soils_all$high[ohio_soils_all$year == 2022], accuracy = 5)`.
    - From 2019-2022 a `r percent((ohio_soils_all$expected[ohio_soils_all$year == 2022] - ohio_soils_all$expected[ohio_soils_all$year == 2019]) / ohio_soils_all$expected[ohio_soils_all$year == 2019], accuracy = 1)` change from the previous triennial update.

## Projected CAUV Expectations

- Non-land costs have trended downward since 2015 and since this component is based off of an Olympic Average of the past 7 years, high values for non-land costs will be dropping out and this component is expected to continue to decrease.
- Prices for corn, soybeans, and wheat have stagnated and are projected at current levels. Component based off of an Olympic Average of the past 8 years (excludes current). High values from 2012 and 2013 are dropping out.
- Yields are based off of the past 11 years (excludes current) and have mostly continued along the path of a 30-year trendline upward.
- Capitalization rate is based on an Olympic Average of the past 7 years which has recently been trending downward. Current Federal Reserve policy is pushing downward pressure on capitalization rate into the future.

## Expectations

```{r cauv-expected, fig.cap="\\label{fig:cauv-expected}"}
```

## Expectations

```{r cauv-expected-table}
```

## Updates

```{r update-map, fig.cap="\\label{fig:update-map}"}
```


# CAUV Calculation

## CAUV Calculation Overview

> It is the expected net present value of an acre of land based. To determine this, first a historical average of yields and prices for corn, soybeans, and wheat to determine gross income. Then historical non-land costs – compliments of The Ohio State University Agricultural Extension Agency – are subtracted from gross income for a measure of net income. And finally, this net income is divided by a capitalization rate based upon historical values of farm interest and equity rates. This value will vary based upon the particular soil type(s) for a farm.

## CAUV Calculation Overview

- A CAUV value is assigned for each of the 3,514 soil types in Ohio
    - based on their expected yields of corn, soybeans, and wheat from the last state-wide soil survey in 1984
- Value is the expected net present value based on the past 5 to 10 years worth of data for each component
    -  Economics jargon, it is the net operating income divided by **capitalization rate**
- Net operating income is based off of the expected revenue (**price** times **yield**) less the expected costs (only **non-land costs**). Calculated for corn, soybeans, and wheat.
- **Capitalization rate** is a combination of the interest rate on a loan for purchasing farmland along with appreciation rate of farmland.

## Generic Formula

- CAUV is assigned for each soil type, although what really matters is the yields that soil type had for corn, soybeans, and wheat in 1984:

$$
CAUV_{s,t} = \frac{NOI_{s,t}}{CAP_t}
$$

- where $CAP_t$ represents the capitalization rate and $NOI_{s,t}$ represents the net operating income based on revenues less non-land costs for corn, soybeans, and wheat.
- Net operating income, ${NOI_{s,t}}$, captures the average returns to an acre of land under normal management practices which is adjusted by the state-wide rotation pattern of commodities.

## Generic Formula

- ${NOI_{s,t}}$ can be defined as:

$$
NOI_{s,t} = \sum_{c} w_{c,t}\times(GOI_{s,c,t} - {nonland}_{s,c,t})
$$

- where $w_{c,t}$ represents a commodities share of total harvested acreage
- $GOI_{s,c,t}$ is the expected revenues for a soil type depending on the statewide price of a commodity and yield is a percentage adjustment to the soil's yield in 1984 for each commodity:

$$
GOI_{s,c,t} = Price_{c,Ohio,t} \times \left( \frac{Yield_{c,Ohio,t}}{Yield_{c,Ohio,1984}} \times Yield_{c,s,1984} \right)
$$


# Component Trends

## Non-Land Costs

```{r viz-nonland, fig.cap="\\label{fig:viz-nonland}"}
```

## Prices

```{r viz-prices, fig.cap="\\label{fig:prices}"}
```

## Yields

```{r viz-yields, fig.cap="\\label{fig:yields}"}
```

## Capitalization Rate

```{r viz-cap, fig.cap="\\label{fig:viz-cap}"}
```


# Component Values

## Rotations


```{r corn-rot}
```

---

```{r soy-rot}
```

---

```{r wheat-rot}
```

## Non-Land Costs


```{r corn-base}
```

---

```{r corn-add}
```

---

```{r soy-base}
```

---

```{r soy-add}
```

---

```{r wheat-base}
```

---

```{r wheat-add}
```

## Prices

```{r corn-price}
```

---

```{r soy-price}
```

---

```{r wheat-price}
```

## Yields

```{r corn-yield}
```

---

```{r soy-yield}
```

---

```{r wheat-yield}
```

## Capitalization Rate


```{r cap-table}
```


# CAUV Values

## Expectation

```{r cauv-expected, fig.cap="\\label{fig:cauv-expected2}"}
```

---

```{r cauv-expected-table}
```

## High Range

```{r cauv-high, fig.cap="\\label{fig:cauv-high}"}
```

---

```{r cauv-high-table}
```

## Low Range

```{r cauv-low, fig.cap="\\label{fig:cauv-low}"}
```

---

```{r cauv-low-table}
```


# References

---

- Farm Office annual crop budget reports https://farmoffice.osu.edu/farm-management-tools/farm-budgets
- Ohio Code of Legislation http://codes.ohio.gov/orc/5713.31 and http://codes.ohio.gov/orc/5715.01
- ODT CAUV Information page https://www.tax.ohio.gov/real_property/cauv.aspx
- USDA-NASS price and yield data https://quickstats.nass.usda.gov
- USDA-ERS total rate of return on farm equity data https://www.ers.usda.gov/data-products/farm-income-and-wealth-statistics/data-files-us-and-state-level-farm-income-and-wealth-statistics/
- Projections for 2020 CAUV values are available at https://github.com/rdinter/CAUV/blob/master/3-proj/future/projections_tidy_2020.csv
- Projections for 2021 CAUV values are available at https://github.com/rdinter/CAUV/blob/master/3-proj/future/projections_tidy_2021.csv
- Projections for 2022 CAUV values are available at https://github.com/rdinter/CAUV/blob/master/3-proj/future/projections_tidy_2022.csv
